Beta Diversity through Time and Space

### 2020-09-10 analyses/beta_diversity_v2.Rmd
haaaands

### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
```{r}
### Elevation diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  theme_abyss()

### Time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  theme_abyss()

### Elevation diff within 1 wk time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                metric.class != "OS.partition" &
                yday.diff <= 7), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.15, pch = 16, size = 1) +
  geom_smooth() +
  facet_wrap(~metric.class) +
  theme_light()

### Time diff within 250 m elev diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                metric.class != "OS.partition" &
                elev.diff <= 250), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.15, pch = 16, size = 1) +
  geom_smooth() +
  facet_wrap(~metric.class) +
  theme_light()

### 2D ElevationDiff*TimeDiff
gam_betadiv <- bamV(value ~ te(elev.diff, yday.diff, 
                               bs = c("te", "te"),
                               k = c(10, 10)),
                    family = "quasibinomial",
                    data = filter(beta_mit_data, year1 == year2 &
                                    metric == "WN"),
                    method = "REML")

gam_betadiv <- bam(value ~ 
                     s(year1, bs = "re") +
                     te(yday.diff, elev.diff, k = c(10, 10)),
                   data = filter(beta_mit_data, year1 == year2 &
                                   metric == "WN.rich"),
                   family = "quasibinomial",
                   discrete = TRUE,
                   method = "fREML") %>% getViz()

check.gamViz(gam_betadiv)
plot(gam_betadiv)

ggplot(filter(beta_site.date, year1 == year2 & metric == "WN"),
       aes(value)) +
  geom_histogram()
```

>>>>

### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
```{r}
### Elevation diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition")), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_brewer(type = "qual", palette = 6) +
  theme_abyss()

### Time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition")), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.25, pch = 16) +
  geom_smooth() +
  facet_grid(metric.class ~ year1) +
  scale_color_brewer(type = "qual", palette = 6) +
  theme_abyss()

### Elevation diff within 1 wk time diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition") &
                yday.diff <= 7), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.1, pch = 16, size = 1) +
  geom_smooth() +
  facet_wrap(~metric.class) +
  facet_grid(metric.class ~ year1) +
  scale_color_brewer(type = "qual", palette = 6) +
  theme_light()

### Time diff within 250 m elev diff
ggplot(filter(beta_mit_data, year1 == year2 &
                site1 != site2 &
                metric != "S" & 
                !metric.class %in% c("OS.partition", "WN.partition") &
                elev.diff <= 250), 
       aes(yday.diff, value, color = metric)) +
  geom_point(alpha = 0.1, pch = 16, size = 1) +
  geom_smooth() +
  facet_wrap(~metric.class) +
  scale_color_brewer(type = "qual", palette = 6) +
  theme_light()
```
